The effect of restricted substrate diffusion on microbial community function and structure

### 2019-03-11 SI_code.Rmd
...

---
title: Variable Composition of Microbial Biomass Affects the Accuracy of Microbial
  Explicit Models Simulation
author: "Petr Capek, Katherine Todd-Brown, Natalie Sadler, Nancy Hess, Kirsten Hofmockel"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

@@

##libraries
library(deSolve)
library(dplyr)
library(FME)
library(reshape)
library(ggplot2)
library(foreach)
library(doParallel)
library(openxlsx)
library(DEoptim)
library(gridExtra)
library(kableExtra)
library(lmerTest)
library(psycho)
library(ggeffects)
library(rcompanion)

@@

@@

t.s0<-as.data.frame(cbind(anova(lme1)[c(5,3,6)], anova(lme3)[c(5,3,6)], anova(lme3)[c(5,3,6)]))

@@

kable(t.s0, digits = c(2, 1, 3, 2, 1, 3, 2, 1, 3), align = 'c', "html", booktabs=T, col.names = c("F", "df", "p", "F", "df", "p", "F", "df", "p"), row.names=T, caption = "_**Table S1:** Results of linear mixed-effects model analysis. Effect of substrate, the type physical barrier (Structure) and their interaction was evaluated. Statistical analysis was conducted for measured respiration rate, cellular protien concentration and DNA. F values, degrees of freedom (df) and the probabality (p) are reported._") %>% 
  kable_styling() %>% 
  add_header_above(c(" " = 1, "Respiration rate" = 3, "Cellular Proteins" = 3, "DNA" = 3),
                   italic=T)

@@

@@

@@

kable(t.s1.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$'), row.names=T, caption = "_**Table S2:** Monod model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monod_r_results_good, digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S3:** Goodness of fit of monod model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 2) %>%
  group_rows("Substrate", 3, 4) %>%
  group_rows("Each", 4, 4)

@@

kable(t.s2.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$'), row.names=T, caption = "_**Table S4:** MEND model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(mend_r_results_good, digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S5:** Goodness of fit of MEND model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 2) %>%
  group_rows("Substrate", 3, 4) %>%
  group_rows("Each", 4, 4)

@@

kable(t.s3.1, digits = c(2, 2, 3, 2, 2), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', '$\\m_{R}$', '$\\f_{0}$', '$\\Y_{S}$'), row.names=T, caption = "_**Table S6:** DEB model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(deb_r_results_good, digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S7:** Goodness of fit of DEB model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 2) %>%
  group_rows("Substrate", 3, 4) %>%
  group_rows("Each", 4, 4)

@@

kable(t.s4.1, digits = c(0, 0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" "," ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$', 'Conversion factor'), row.names=F, caption = "_**Table S8:** Monod model parameters calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 2) %>%
  group_rows("Structure", 3, 8) %>%
  group_rows("Substrate", 9, 12) %>%
  group_rows("Each", 13, 24)

kable(monod_all_results_good, digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S9:** Goodness of fit of Monod model calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("Protein", 1, 8, label_row_css = "background-color: #666; color: #fff;") %>%
  group_rows("All", 1, 2) %>%
  group_rows("Structure", 3, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 8) %>%
  group_rows("DNA", 9, 16, label_row_css = "background-color: #666; color: #fff;") %>%
  group_rows("All", 9, 10) %>%
  group_rows("Structure",11, 12) %>%
  group_rows("Substrate", 13, 14) %>%
  group_rows("Each", 15, 16)

@@

kable(t.s5.1, digits = c(0, 0, 2, 2, 2, 3, 2), align = 'c', "html", booktabs=T, col.names = c(" "," ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$', 'Conversion factor'), row.names=F, caption = "_**Table S10:** MEND model parameters calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 2) %>%
  group_rows("Structure", 3, 8) %>%
  group_rows("Substrate", 9, 12) %>%
  group_rows("Each", 13, 24)

kable(mend_all_results_good, digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S11:** Goodness of fit of MEND model calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("Protein", 1, 8, label_row_css = "background-color: #666; color: #fff;") %>%
  group_rows("All", 1, 2) %>%
  group_rows("Structure", 3, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 8) %>%
  group_rows("DNA", 9, 16, label_row_css = "background-color: #666; color: #fff;") %>%
  group_rows("All", 9, 10) %>%
  group_rows("Structure",11, 12) %>%
  group_rows("Substrate", 13, 14) %>%
  group_rows("Each", 15, 16)

@@

kable(t.s6.1, digits = c(0, 2, 2, 3, 2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" ",'$\\V_{MAX}$', '$\\K_{M}$', '$\\m_{R}$', '$\\f_{0}$', '$\\Y_{S}$', '$\\R_{Proteins}$', '$\\S_{Proteins}$', '$\\S_{DNA}$'), row.names=F, caption = "_**Table S12:** DEB model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(deb_all_results_good, digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S13:** Goodness of fit of DEB model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 3) %>%
  group_rows("Structure", 4, 6) %>%
  group_rows("Substrate", 7, 9) %>%
  group_rows("Each", 10, 12)

@@

>>>>

---
title: The effect of restricted substrate diffusion on microbial community function and structure
author: "Petr Capek, Katherine Todd-Brown, Natalie Sadler, Nancy Hess, Kirsten Hofmockel"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

@@

##libraries
library(deSolve)
library(dplyr)
library(FME)
library(reshape)
library(ggplot2)
library(foreach)
library(doParallel)
library(openxlsx)
library(DEoptim)
library(gridExtra)
library(kableExtra)
library(lmerTest)
library(psycho)
library(ggeffects)
library(rcompanion)
library(vegan)

@@

#Bacteria
bac<-as.data.frame(t(read.csv(file=c("bacteria.csv"), skip = 4)[,-1]))
colnames(bac)<-t(read.csv(file=c("bacteria.csv"), skip = 4)[,1])

bac_env<-as.data.frame(t(read.csv(file=c("bacteria.csv"), nrows = 4, header = F)[,-1]))
colnames(bac_env)<-t(read.csv(file=c("bacteria.csv"), nrows = 4, header = F)[,1])

bac_env$Structure<-recode(bac_env$Structure, 'Broth'="BROTH", 'Glass wool'="WOOL",
                          'Mixed glass'="GLASS")

#Fungi
fungi<-as.data.frame(t(read.csv(file=c("fungi.csv"), skip = 4)[,-1]))
colnames(fungi)<-t(read.csv(file=c("fungi.csv"), skip = 4)[,1])

fungi_env<-as.data.frame(t(read.csv(file=c("fungi.csv"), nrows = 4, header = F)[,-1]))
colnames(fungi_env)<-t(read.csv(file=c("fungi.csv"), nrows = 4, header = F)[,1])

fungi_env$Structure<-recode(fungi_env$Structure, 'Broth'="BROTH", 'Glass wool'="WOOL",
                          'Mixed glass'="GLASS")

@@

t.s0<-as.data.frame(cbind(anova(lme1)[c(5,3,6)], anova(lme2)[c(5,3,6)], anova(lme3)[c(5,3,6)]))

@@

kable(t.s0, digits = c(2, 1, 3, 2, 1, 3, 2, 1, 3), align = 'c', "html", booktabs=T, col.names = c("F", "df", "p", "F", "df", "p", "F", "df", "p"), row.names=T, caption = "_**Table S1:** Results of linear mixed-effects model analysis. Effect of substrate, the type of physical barrier (Structure) and their interaction was evaluated. Statistical analysis was conducted for measured respiration rate, cellular protein concentration and DNA. F values, degrees of freedom (df) and the probability (p) are reported._") %>% 
  kable_styling() %>% 
  add_header_above(c(" " = 1, "Respiration rate" = 3, "Cellular Proteins" = 3, "DNA" = 3),
                   italic=T)

@@

```
$~$
$~$
$~$
Next, the effect of the substrate and the presence of a physical barrier on microbial community structure and diversity is evaluated. First, the jaccard distance indices are calculated on a normalized data. Then, Permutational Multivariate Analysis of Variance is used to estimate the treatment effects. Analysis permutations are constrained to an individual times of incubation to avoid the potential confounding effect of time.

```{r vegan analysis, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#Bacteria
bac.norm<-decostand(bac, method=c("normalize"))
bac.dist<-vegdist(bac.norm, method = "jaccard")
bac.ad<-adonis(bac.dist~Substrate+Structure/Substrate, bac_env, strata = bac_env$Day)

#Fungi
fungi.norm<-decostand(fungi, method=c("normalize"))
fungi.dist<-vegdist(fungi.norm, method = "jaccard")
fungi.ad<-adonis(fungi.dist~Substrate+Structure/Substrate, fungi_env, strata = fungi_env$Day)

t.s01<-cbind(as.data.frame(bac.ad$aov[c(4, 1, 6)])[c(1:3), ], as.data.frame(fungi.ad$aov[c(4, 1, 6)])[c(1:3), ])

row.names(t.s01)<-c("Substrate", "Structure", "Substrate*Structure")

kable(t.s01, digits = c(2, 1, 3, 2, 1, 3), align = 'c', "html", booktabs=T, col.names = c("F", "df", "p", "F", "df", "p"), row.names=T, caption = "_**Table S2:** Results of permutational multivariate analysis of variance. Effect of substrate, the type of physical barrier (Structure) and their interaction was evaluated. Statistical analysis was conducted for bacterial and fungal part of the microbial community separately. F values, degrees of freedom (df) and the probability (p) are reported._") %>% 
  kable_styling() %>% 
  add_header_above(c(" " = 1, "Bacteria" = 3, "Fungi" = 3),
                   italic=T)

```
$~$
$~$
$~$
$~$
$~$
$~$
```{r community figure, echo=FALSE, fig.height=5, fig.width=10, fig.cap="Fig. S2: Multivariate distances between six experimental groups that combine two different substrate treatments (Glucose or Cellobiose) and three different types of physical barrier (BROTH, WOOL, GLASS - black, grey and empty symbols respectively) in respect to structure of bacterial ((a); 16S) and fungal ((b); ITS) part of microbial community."}
#Bacteria
bacm<-as.matrix(bac.dist)
#Intercept - Glucose - BROTH
mGB<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
       intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Glucose"))]
mGB[mGB==0]<-NA
#Glucose - GLASS
mGG<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
       intersect(which(bac_env$Structure=="GLASS"),which(bac_env$Substrate=="Glucose"))]
mGG[mGG==0]<-NA
#Glucose - WOOL
mGW<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
       intersect(which(bac_env$Structure=="WOOL"),which(bac_env$Substrate=="Glucose"))]
mGW[mGW==0]<-NA
#Cellobiose - BROTH
mCB<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
       intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose"))]
mCB[mCB==0]<-NA
#Cellobiose - GLASS
mCG<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
       intersect(which(bac_env$Structure=="GLASS"),which(bac_env$Substrate=="Cellobiose"))]
mCG[mCG==0]<-NA
#Cellobiose - WOOL
mCW<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
       intersect(which(bac_env$Structure=="WOOL"),which(bac_env$Substrate=="Cellobiose"))]
mCW[mCW==0]<-NA

bacdiff<-data.frame(Dist=c(mean(mGB, na.rm=T), mean(mGG, na.rm=T), mean(mGW, na.rm=T), 0, mean(mCG, na.rm=T), mean(mCW, na.rm=T)),
                    Dist.sd=c(sd(mGB, na.rm=T), sd(mGG, na.rm=T), sd(mGW, na.rm=T), sd(mCB, na.rm=T), sd(mCG, na.rm=T), sd(mCW, na.rm=T)),
                    Substrate=c(rep("Glucose", 3), rep("Cellobiose", 3)),
                    Structure=rep(c("BROTH", "GLASS", "WOOL"), 2))

#Fungi
fm<-as.matrix(fungi.dist)
#Intercept - Glucose - BROTH
fGB<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
          intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Glucose"))]
fGB[fGB==0]<-NA
#Glucose - GLASS
fGG<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
          intersect(which(fungi_env$Structure=="GLASS"),which(fungi_env$Substrate=="Glucose"))]
fGG[fGG==0]<-NA
#Glucose - WOOL
fGW<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
          intersect(which(fungi_env$Structure=="WOOL"),which(fungi_env$Substrate=="Glucose"))]
fGW[fGW==0]<-NA
#Cellobiose - BROTH
fCB<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
          intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose"))]
fCB[fCB==0]<-NA
#Cellobiose - GLASS
fCG<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
          intersect(which(fungi_env$Structure=="GLASS"),which(fungi_env$Substrate=="Cellobiose"))]
fCG[fCG==0]<-NA
#Cellobiose - WOOL
fCW<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
          intersect(which(fungi_env$Structure=="WOOL"),which(fungi_env$Substrate=="Cellobiose"))]
fCW[fCW==0]<-NA

fdiff<-data.frame(Dist=c(mean(fGB, na.rm=T), mean(fGG, na.rm=T), mean(fGW, na.rm=T), 0, mean(fCG, na.rm=T), mean(fCW, na.rm=T)),
                    Dist.sd=c(sd(fGB, na.rm=T), sd(fGG, na.rm=T), sd(fGW, na.rm=T), sd(fCB, na.rm=T), sd(fCG, na.rm=T), sd(fCW, na.rm=T)),
                    Substrate=c(rep("Glucose", 3), rep("Cellobiose", 3)),
                    Structure=rep(c("BROTH", "GLASS", "WOOL"), 2))

bacplot<-ggplot(bacdiff, aes(Substrate, Dist))+
  geom_point(position = position_dodge(width = 0.5), aes(fill=Structure), shape=21, cex=6)+
  scale_fill_manual(values = c("black", "grey", "white"))+
  scale_colour_manual(values = c("black", "black", "black"))+
  geom_errorbar(aes(ymin=Dist-Dist.sd, ymax=Dist+Dist.sd, colour=Structure),
                position=position_dodge(width = 0.5), width=0.1)+
  ylab("16S Distance")+
  ggtitle("a)")+
  theme_min+theme(axis.title.x = element_blank(),
                  legend.title = element_blank(),
                  legend.position = c(0.8, 0.2))
fplot<-ggplot(fdiff, aes(Substrate, Dist))+
  geom_point(position = position_dodge(width = 0.5), aes(fill=Structure), shape=21, cex=6, show.legend = F)+
  scale_fill_manual(values = c("black", "grey", "white"))+
  scale_colour_manual(values = c("black", "black", "black"))+
  geom_errorbar(aes(ymin=Dist-Dist.sd, ymax=Dist+Dist.sd, colour=Structure),
                position=position_dodge(width = 0.5), width=0.1, show.legend = F)+
  ylab("ITS Distance")+
  ggtitle("b)")+
  theme_min+theme(axis.title.x = element_blank(),
                  legend.title = element_blank())

grid.arrange(bacplot, fplot, nrow=1)

@@

$~$
$~$
$~$

@@

kable(t.s1.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$'), row.names=T, caption = "_**Table S3:** Monod model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monod_r_results_good, digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S4:** Goodness of fit of monod model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 2) %>%
  group_rows("Substrate", 3, 4) %>%
  group_rows("Each", 4, 4)

@@

kable(t.s2.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$'), row.names=T, caption = "_**Table S5:** MEND model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(mend_r_results_good, digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S6:** Goodness of fit of MEND model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 2) %>%
  group_rows("Substrate", 3, 4) %>%
  group_rows("Each", 4, 4)

@@

kable(t.s3.1, digits = c(2, 2, 3, 2, 2), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', '$\\m_{R}$', '$\\f_{0}$', '$\\Y_{S}$'), row.names=T, caption = "_**Table S7:** DEB model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(deb_r_results_good, digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S8:** Goodness of fit of DEB model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 2) %>%
  group_rows("Substrate", 3, 4) %>%
  group_rows("Each", 4, 4)

@@

kable(t.s4.1, digits = c(0, 0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" "," ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$', 'Conversion factor'), row.names=F, caption = "_**Table S9:** Monod model parameters calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 2) %>%
  group_rows("Structure", 3, 8) %>%
  group_rows("Substrate", 9, 12) %>%
  group_rows("Each", 13, 24)

kable(monod_all_results_good, digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S10:** Goodness of fit of Monod model calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("Protein", 1, 8, label_row_css = "background-color: #666; color: #fff;") %>%
  group_rows("All", 1, 2) %>%
  group_rows("Structure", 3, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 8) %>%
  group_rows("DNA", 9, 16, label_row_css = "background-color: #666; color: #fff;") %>%
  group_rows("All", 9, 10) %>%
  group_rows("Structure",11, 12) %>%
  group_rows("Substrate", 13, 14) %>%
  group_rows("Each", 15, 16)

@@

kable(t.s5.1, digits = c(0, 0, 2, 2, 2, 3, 2), align = 'c', "html", booktabs=T, col.names = c(" "," ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$', 'Conversion factor'), row.names=F, caption = "_**Table S11:** MEND model parameters calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 2) %>%
  group_rows("Structure", 3, 8) %>%
  group_rows("Substrate", 9, 12) %>%
  group_rows("Each", 13, 24)

kable(mend_all_results_good, digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S12:** Goodness of fit of MEND model calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("Protein", 1, 8, label_row_css = "background-color: #666; color: #fff;") %>%
  group_rows("All", 1, 2) %>%
  group_rows("Structure", 3, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 8) %>%
  group_rows("DNA", 9, 16, label_row_css = "background-color: #666; color: #fff;") %>%
  group_rows("All", 9, 10) %>%
  group_rows("Structure",11, 12) %>%
  group_rows("Substrate", 13, 14) %>%
  group_rows("Each", 15, 16)

@@

kable(t.s6.1, digits = c(0, 2, 2, 3, 2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" ",'$\\V_{MAX}$', '$\\K_{M}$', '$\\m_{R}$', '$\\f_{0}$', '$\\Y_{S}$', '$\\R_{Proteins}$', '$\\S_{Proteins}$', '$\\S_{DNA}$'), row.names=F, caption = "_**Table S13:** DEB model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(deb_all_results_good, digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S14:** Goodness of fit of DEB model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 3) %>%
  group_rows("Structure", 4, 6) %>%
  group_rows("Substrate", 7, 9) %>%
  group_rows("Each", 10, 12)

@@

$~$
$~$
$~$
$~$
$~$
$~$
```{r correlation figure, echo=FALSE, fig.height=5, fig.width=10, fig.cap="Fig. S3: Effect of structure of bacterial and fungal part of microbial community (expressed as multivariate mean distance between groups; see Matrial and Methods section) on overall microbial community maintnance respiration rate ($m_{R}$ - (a)) and protein concentration in Reserves pool ($R_{Protein}$ - (b)). The value of both parameters are expressed as a mean difference between groups in order to allow unbiased correlation with microbial community structure metric (see Matrial and Methods section)."}

dpars<-t.s6.1[c(7:12), -1]
dpars$Substrate<-rep(c("Cellobiose", "Glucose"), 3)
dpars$Structure<-c("BROTH", "BROTH", "GLASS", "GLASS", "WOOL", "WOOL")

dpars2<-merge(dpars, bacdiff, by.y=c("Substrate", "Structure"), by.x=c("Substrate", "Structure"))

dpars3<-merge(dpars, fdiff, by.y=c("Substrate", "Structure"), by.x=c("Substrate", "Structure"))

for(i in 3:10){
  dpars2[, i]<-sqrt((dpars2[, i]-as.numeric(dpars2[1, i]))^2)
}

for(i in 3:10){
  dpars3[, i]<-sqrt((dpars3[, i]-as.numeric(dpars3[1, i]))^2)
}

dpars2$group<-c("Bacteria")
dpars3$group<-c("Fungi")
dpars4<-rbind(dpars2, dpars3)

corplot1.p<-coefficients(nls(m0~Dist^x, dpars4, start = list(x=2)))

corplot1<-ggplot(dpars4[dpars4$group=="Bacteria", ], aes(Dist, m0*1e4))+
  geom_point(cex=6, aes(fill=Structure), show.legend=T, shape=21)+
  geom_point(data=dpars4[dpars4$group=="Fungi", ], cex=6, 
             aes(x=Dist, y=m0*1e4, fill=Structure), show.legend=F, shape=22)+
  scale_fill_manual(values = c("black", "grey", "white"))+
  stat_function(fun = function(x){(x^corplot1.p)*1e4}, lwd=1, colour="grey30")+
  xlab("16S and ITS Distance")+
  ggtitle("a)")+
  ylab(expression(paste(m[R]~difference)))+
  theme_min+theme(legend.title = element_blank(),
                  legend.position = c(0.2,0.8))

corplot2.p<-coefficients(nls(fpr~Dist^x, dpars4, start = list(x=2)))

corplot2<-ggplot(dpars4[dpars4$group=="Bacteria", ], aes(Dist, fpr))+
  geom_point(cex=6, aes(fill=Structure), show.legend=F, shape=21)+
  geom_point(data=dpars4[dpars4$group=="Fungi", ], cex=6, 
             aes(x=Dist, y=fpr, fill=Structure), show.legend=F, shape=22)+
  scale_fill_manual(values = c("black", "grey", "white"))+
  stat_function(fun = function(x){(x^corplot2.p)}, lwd=1, colour="grey30")+
  xlab("16S and ITS Distance")+
  ggtitle("b)")+
  ylab(expression(paste(R[Proteins]~difference)))+
  theme_min

grid.arrange(corplot1, corplot2, nrow=1)
